Spin canonical divisors of log stacky curves

### 2015-06-27 stacky_write_up/spin_cring.tex
fix delta-not-1 typo
finished g=2 base

We wish to show $f$ is surjective.
Note that $\dim_k \ker f = \dim_k H^0((n-2)L) = (n-3)(g - 1)$ using Riemann--Roch and the assumption $n \geq 5$ or ($n = 4$ and $\Delta > 0$). Note that here we are implicitly using the Remark ~\ref{rem:delta-not-1}
Additionally, $\dim_k V \otimes H^0(nL) = 2 \cdot (n- 1)(g - 1)$, again using Riemann--Roch.
Therefore, $$\dim_k \im f = 2 \cdot (n- 1)(g - 1) -(n-3)(g - 1) = (n+ 1)(g - 1) = \dim_k H^0((n+2)L),$$
again using Riemann-Roch, implying $f$ is surjective.
\end{proof}

\begin{prop}
\label{prop:genus_2_generation_5}
Let $X$ be a smooth genus 2 curve with canonical divisor $K$ and let $L \in \di X$ be so that $2L \sim K$. 
Then, $R_L$ is generated in degree at most $5$.
\end{prop}
\begin{proof}
By Lemma ~\ref{lem:semicanonical_generation} it suffices to show there are no new generators in degree 6.
Let $P_1, \ldots, P_6$ be the six hyperelliptic fixed points of $X$. 
Recall that all spin divisors on $X$ can be written as 
$$L = P_1 + \sum_{i =2}^{6} \epsilon_i (P_1 - P_i),$$ 
with $\epsilon_i \in \{0,1\},$ and some $i \in \{2, 3, 4, 5, 6\}$ for which $\epsilon_i = 0$. \todo{cite this somewhere, maybe ask david for a place to cite?}
Note that $H^0(X,K)$ is spanned by rational functions $u, x$ with $\di u|_{2L} = -2L, \di x|_{2L} = -2L + 2P_1.$ Therefore, the image of the multiplication map 
\begin{align*}
	Sym^3 H^0(X,2L) \rightarrow H^0(X,6L)
\end{align*}
is 4 dimensional by Riemann--Roch, with image spanned by $u^3,u^2x,ux^2, x^3$, 
which have $\di u^3|_{2L} = -2L, \di u^2x|_{2L} = -2L + 2P_1,\di ux^2|_{2L} = -2L+4P_1, \di x^3|_{2L} = -2L + 6P_1.$ 
Since $H^0(X,6L)$ is 5 dimensional by Riemann--Roch,
it suffices to produce one more independent function in the image of the multiplication map.
There are now two cases, depending on whether $H^0(X,L) = 0$.

First, we deal with the case that $H^0(X,L) \neq 0$. 
Then, since $X \not \cong \BP^1_k,$ and $\deg L = 1$, we must have $H^0(X,L) = 1$. Hence, we know $L$ is effective, so we must have $L = P$ where $P$ is hyperelliptic fixed . In this case, it is shown in \cite{neves:halfcan}[Theorem III.19] that $R_L$ is generated in degree $\leq 3$.

@@

\todo{This should be $k[x_1, \ldots, x_n]$ instead of $\sym H^0$.}
\begin{lem}
\label{lem:reducing_degree}
Suppose $X$ is a curve of genus $g \geq 2$ and let $L$ be a divisor with $2L \sim K+\Delta$. Choose generators $x_1, \ldots, x_n$ of $R_L$ so that we obtain a surjection $\phi:k[x_1, \ldots, x_n] \twoheadrightarrow R_L$ with kernel $I_L$. Let $I_{L,d}$ be the $d$th graded piece of $I_L$ and let $J_{L,d} = \sum_{j = 1}^{d- 1}k[x_1, \ldots, x_n]_j \cdot I_{L,d-j}.$ For any $f \in k[x_1, \ldots, x_n]$ such that
$$
\deg f \geq \begin{cases}
	12 &\text{ if }\Delta = 0\\
	11 &\text{ if }\Delta > 0
\end{cases}$$
there exist $s_1,s_2 \in k[x_1, \ldots, x_n]_2$ and $g,h \in k[x_1, \ldots, x_n]_{d-2}$ so that $s_1g + s_2h = f \bmod J_{L,d}.$
\end{lem}
\begin{proof}
Choose $s_1,s_2 \in k[x_1, \ldots, x_n]_2$ so that $\phi(s_1),\phi(s_2) \in H^0(2L)$ are independent elements.
By Lemma ~\ref{lem:generation_5}, $\deg x_i \leq 5$ for $1 \leq i \leq n$. Then, we may write $f = \sum_{i = 1}^{n}a_i x_i$ with $a_i \in k[x_1, \ldots, x_n]_{d-\deg x_i}$.
We will next show that for all $1 \leq i \leq n$ there exist $g_i, h_i \in k[x_1, \ldots, x_n]_{d - \deg x_i - 2}$ so that $a_i = s_1g_i + s_2h_i \bmod I_{L,\deg a_i}.$ 
Let $V \subset H^0(K)$ be the vector subspace generated by $\phi(s_1)$ and $\phi(s_2)$. By Lemma ~\ref{lem:semicanonical_generation},
\begin{align*}
	V \otimes H^0((\deg f-\deg x_i -2)L) \rightarrow H^0((\deg f-\deg x_i)L)
\end{align*}
is surjective, since
$$
\deg f- \deg a_i -2 \geq \begin{cases}
	5 &\text{ if }\Delta = 0\\
	4 &\text{ if }\Delta > 0.
\end{cases}$$
In particular, we can write $\phi(a_i) = \phi(s_1) \alpha + \phi(s_1) \cdot \beta$. Choosing $g_i,h_i$ so that $\phi(g_i) = \alpha,\phi(h_i) = \beta$, we have $a_i = s_1 g_i + s_2 h_i \bmod I_{L,\deg a_i}$, as claimed.

@@

\begin{prop}
\label{prop:relation_11}
Let $X$ be a curve of genus $g \geq 2$ and let $L$ be a divisor with $2L \sim K + \Delta$. Then $I_L$ is generated in degree at most $11$ if $\Delta = 0$ and degree at most $10$ if $\Delta > 0$.
\end{prop}
\begin{proof}
Suppose $f \in I_L$ with $\deg f \geq 12$. To complete the proof, it suffices to show $f \in I_{L,d}'$. By Lemma ~\ref{lem:reducing_degree}, it suffices to show $s_1g+s_2h \in I_{L,d}'$. 
Let $V \subset H^0(K)$ be the vector subspace generated by $\phi(s_1)$ and $\phi(s_2)$, and consider the map
$$\begin{tikzcd}
V \otimes H^0((\deg f - 2)L) \ar {r}{f} & H^0((\deg f)L),
\end{tikzcd}$$
we know that $\phi(s_1)\phi(g) + \phi(s_2) \phi(h) \mapsto 0.$
So, using the refined version of the basepoint-free pencil trick, as shown in the proof of \cite{saint-donat:proj}[Lemma 2.6], if $\phi(s_1) g + \phi(s_2)h \mapsto 0,$ there exists some $\rho \in k[x_1, \ldots, x_n]$ so that $\phi(\rho) \in H^0((\deg f - 4)L)$ satisfies $\phi(g) = \phi(s_2)\phi(\rho)$ and $\phi(h) = -\phi(s_1)\phi(\rho).$ Therefore, $g \equiv s_2 \rho \bmod I_{k,d-2}$ and $h \equiv -s_1 \rho \bmod I_{k,d-2}$. Therefore,
\begin{align*}
	s_1g + s_2h \equiv s_1s_2\rho + s_2(-s_1 \rho) \equiv 0 \bmod J_{L,d}.
\end{align*}
\end{proof}

\begin{rem}
\label{rem:relations_generation_ten}
In \cite{neves:halfcan}[Theorem III.19], Neves shows that if $L=P$ is a half canonical divisor on a genus 2 curve, then $R_L$ has a relation in degree 10. Therefore, the bound from Proposition ~\ref{prop:relation_11}, that $I_L$ is generated in degree at most 11 is close to sharp.
\end{rem}

>>>>

We wish to show $f$ is surjective.
Note that $\dim_k \ker f = \dim_k H^0((n-2)L) = (n-3)(g - 1)$ using Riemann--Roch and the assumption $n \geq 5$ or ($n = 4$ and $\Delta > 0$). Note that here we are implicitly using the Remark ~\ref{rem:delta-not-1}.
Additionally, $\dim_k V \otimes H^0(nL) = 2 \cdot (n- 1)(g - 1)$, again using Riemann--Roch.
Therefore, $$\dim_k \im f = 2 \cdot (n- 1)(g - 1) -(n-3)(g - 1) = (n+ 1)(g - 1) = \dim_k H^0((n+2)L),$$
again using Riemann-Roch. Ergo, $f$ is surjective.
\end{proof}

\begin{prop}
\label{prop:genus_2_generation_5}
Let $X$ be a smooth genus 2 curve with canonical divisor $K$ and let $L \in \di X$ be so that $2L \sim K$. 
Then, $R_L$ is generated in degree at most $5$.
\end{prop}
\begin{proof}
By Lemma ~\ref{lem:semicanonical_generation} it suffices to show $R_L$ has no minimal generators in degree 6.
Let $P_1, \ldots, P_6$ be the six hyperelliptic fixed points of $X$. 
Recall that all linear equivalence classes of spin divisors on $X$ have a representative of the form
$$L = P_1 + \sum_{i =2}^{6} \epsilon_i (P_1 - P_i),$$ 
with $\epsilon_i \in \{0,1\},$ and not all $\epsilon_i = 1$. \todo{cite this somewhere, maybe ask david for a place to cite? We can cite http://www.math.harvard.edu/~ctm/home/text/class/harvard/213b/14/html/home/course/course.pdf, for example, from complex analysis course notes!}
Note that $H^0(X,K)$ is spanned by rational functions $u, x$ with $\di u|_{2L} = -2L, \di x|_{2L} = -2L + 2P_1,$ where $u,x$ are though of as rational functions. Therefore, the image of the multiplication map 
\begin{align*}
	Sym^3 H^0(X,2L) \rightarrow H^0(X,6L)
\end{align*}
is 4 dimensional by Riemann--Roch, with image spanned by $u^3,u^2x,ux^2, x^3$, 
which have 
\begin{align*}
	\di u^3|_{2L} &= -2L \\
	\di u^2x|_{2L} &= -2L + 2P_1 \\
	\di ux^2|_{2L} &= -2L+4P_1, \\
	\di x^3|_{2L} &= -2L + 6P_1.
\end{align*}

\noindent
Since $H^0(X,6L)$ is 5 dimensional by Riemann--Roch,
it suffices to produce one more independent function in the image of the multiplication map.
There are now two cases, depending on whether $H^0(X,L) = 0$.

First, we deal with the case that $H^0(X,L) \neq 0$. 
Then, since $X \not \cong \BP^1_k,$ and $\deg L = 1$, we must have $H^0(X,L) = 1$. Hence, we know $L$ is effective, so up to linear equivalence, we may assume $L = P$ where $P$ is hyperelliptic fixed. In this case, it is shown in \cite{neves:halfcan}[Theorem III.19] that $R_L$ is generated in degree $\leq 3$.

@@

\begin{lem}
\label{lem:reducing_degree}
Suppose $X$ is a curve of genus $g \geq 2$ and let $L$ be a divisor with $2L \sim K+\Delta$. Choose generators $x_1, \ldots, x_n$ of $R_L$ so that we obtain a surjection $\phi:k[x_1, \ldots, x_n] \twoheadrightarrow R_L$ with kernel $I_L$. Let $I_{L,d}$ be the $d$th graded piece of $I_L$ let 
$$J_{L,d} = \sum_{j = 1}^{d- 1}k[x_1, \ldots, x_n]_j \cdot I_{L,d-j},$$
and let $s_1,s_2 \in k[x_1, \ldots, x_n]_2$ be two elements so that $k\phi(s_1) \oplus k\phi(s_2) = V \subset H^0(X,2L)$ is basepoint free.
For any $f \in k[x_1, \ldots, x_n]$ such that
$$
\deg f \geq \begin{cases}
	12 &\text{ if }\Delta = 0\\
	11 &\text{ if }\Delta > 0
\end{cases}$$
there exist $g,h \in k[x_1, \ldots, x_n]_{d-2}$ so that $s_1g + s_2h = f \bmod J_{L,d}.$
\end{lem}
\begin{proof}
Choose $s_1,s_2 \in k[x_1, \ldots, x_n]_2$ so that $\phi(s_1),\phi(s_2) \in H^0(2L)$ generate a base point free pencil $V = k \phi(s_1) \oplus k \phi(s_2)\subset H^0(X,2L)$.
By Lemma ~\ref{lem:generation_5}, $\deg x_i \leq 5$ for $1 \leq i \leq n$. Then, we may write $f = \sum_{i = 1}^{n}a_i x_i$ with $a_i \in k[x_1, \ldots, x_n]_{d-\deg x_i}$.
We will next show that for all $1 \leq i \leq n$ there exist $g_i, h_i \in k[x_1, \ldots, x_n]_{d - \deg x_i - 2}$ so that $a_i = s_1g_i + s_2h_i \bmod I_{L,\deg a_i}.$ 
By Lemma ~\ref{lem:semicanonical_generation},
\begin{align*}
	V \otimes H^0((\deg f-\deg x_i -2)L) \rightarrow H^0((\deg f-\deg x_i)L)
\end{align*}
is surjective, since
$$
\deg f- \deg a_i -2 \geq \begin{cases}
	5 &\text{ if }\Delta = 0\\
	4 &\text{ if }\Delta > 0,
\end{cases}$$
by the assumption on $\deg f$.
In particular, we can write $\phi(a_i) = \phi(s_1) \alpha + \phi(s_1) \cdot \beta$. Choosing $g_i,h_i$ for $1 \leq i \leq n$ so that $\phi(g_i) = \alpha,\phi(h_i) = \beta$, we have $a_i = s_1 g_i + s_2 h_i \bmod I_{L,\deg a_i}$, as claimed.

@@

\begin{prop}
\label{prop:relation_11}
Let $X$ be a curve of genus $g \geq 2$ and let $L$ be a divisor with $2L \sim K + \Delta$. Then $I_L$ is generated in degree at most $11$ if $\Delta = 0$ and degree at most $10$ if $\Delta > 0$.
\end{prop}
\begin{proof}
Suppose $f \in I_L$ with 
$$
\deg f \geq \begin{cases}
	12 &\text{ if }\Delta = 0\\
	11 &\text{ if }\Delta > 0.
\end{cases}$$
To complete the proof, it suffices to show $f \in I_{L,d}'$. By Lemma ~\ref{lem:reducing_degree}, it suffices to show $s_1g+s_2h \in I_{L,d}'$ where $\phi(s_1),\phi(s_2) \in H^0(X,2L)$ are two sections so that $k\phi(s_1) \oplus k \phi(s_2) = V \subset H^0(K)$ is basepoint free. Consider the map
$$\begin{tikzcd}
V \otimes H^0((\deg f - 2)L) \ar {r}{f} & H^0((\deg f)L),
\end{tikzcd}$$
we know that $\phi(s_1)\phi(g) + \phi(s_2) \phi(h) \mapsto 0.$
So, by the explicit isomorphism given in the proof of the basepoint-free pencil trick, as shown in the proof of \cite{saint-donat:proj}[Lemma 2.6], there exists some $\rho \in k[x_1, \ldots, x_n]$ so that $\phi(\rho) \in H^0((\deg f - 4)L)$ satisfies $\phi(g) = \phi(s_2)\phi(\rho)$ and $\phi(h) = -\phi(s_1)\phi(\rho).$ Therefore, $g \equiv s_2 \rho \bmod I_{k,d-2}$ and $h \equiv -s_1 \rho \bmod I_{k,d-2}$. Hence,
\begin{align*}
	s_1g + s_2h \equiv s_1(s_2\rho) + s_2(-s_1 \rho) \equiv 0 \bmod J_{L,d}.
\end{align*}
\end{proof}

\begin{rem}
\label{rem:relations_generation_ten}
In \cite{neves:halfcan}[Theorem III.19], Neves shows that if $L=P$ is a half canonical divisor on a genus 2 curve, then $R_L$ has a relation in degree 10. Therefore, the bound from Proposition ~\ref{prop:relation_11}, that $I_L$ is generated in degree at most 11 if $\Delta = 0$ and at most $10$ if $\Delta > 0$ is close to sharp.
\end{rem}
